---
title: "**NHS Borders** | Cancer Incidence"
author: 
- "John Robert Taylor"
- "[bob.taylor@nhs.net](mailto:bob.taylor@nhs.net)"
- "candidate ID: &nbsp; **20214**"
output:
  pdf_document:
    latex_engine: xelatex
    template: NULL
    keep_tex: false
mainfont: Source Sans Pro
documentclass: proc
urlcolor: blue
geometry:
- layout=a4paper
- top=-10mm
- left=20mm
- right=20mm
- bottom=30mm
subparagraph: yes
header-includes:
- \usepackage{titlesec}
- \titlespacing{\section}{0pt}{2\parskip}{\parskip}
- \titlespacing{\subsection}{0pt}{2\parskip}{\parskip}
- \pagenumbering{gobble}
- \usepackage{titling}
- \setlength{\droptitle}{1.5cm}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gridExtra)
# load plots
p_incidence_trends <- read_rds("data/p_incidence_trends.rds")
p_skin <- read_rds("data/p_skin.rds")
p_skin_network <- read_rds("data/p_skin_network.rds")
p_scot_incidence_demo <- read_rds("data/p_scot_incidence_demo.rds")
p_borders_demo <- read_rds("data/p_borders_demo.rds")
p_both_demos <- read_rds("data/p_both_demos.rds")
```

# Total Cancer Incidence
*NHS Borders* has seen a steady increase in total cancer incidence. In 2017, there were *894* incidences of cancer, compared to *702* one decade earlier -- an increase of over *25%*. This trend is pictured below, from 1993 -- 2017.

```{r, echo=FALSE, fig.width=6, fig.height=3}
p_incidence_trends
```

This increase is mirrored in the crude rate of total cancer incidences per 100,000 population, so is not explicable by any change in the total population numbers.


# Current Outliers & Future Trends

**Selected Cancer Sites:&nbsp;**
In terms of increasing cancer incidence in *NHS Borders*, three cancer sites stand out: *basal cell carcinoma of the skin*; *non-melanoma skin cancer*; *squamos cell carcinoma of the skin*.

```{r, echo=FALSE, fig.width=6, fig.height=3}
p_skin
```

Incidences of these cancers have all roughly doubled in the time period shown, and also exhibit high absolute incidences compared to other cancers in the population of interest. 

It is also worth noting that other geographical regions show similar trends. 

```{r, echo=FALSE, fig.width=6, fig.height=3.5}
p_skin_network
```

Here, the incidence rate (age standardised per 100,000 population, with reference to the European Standard Population) for these cancers is shown, for the three Scottish cancer networks.

**Ageing Population:&nbsp;**
Looking to the future, it is likely that the ageing population will play a role in increasing cancer incidence. In 2017, the median age of an *NHS Scotland* patient diagnosed with cancer was *70* years old. 

```{r, echo=FALSE, fig.width=6, fig.height=3}
p_scot_incidence_demo
```

In *NHS Borders*, population demographics are projected to shift towards the elderly over time. Specifically, the population of over 70s is expected to rise from approx. *20,500* to *30,000* -- an almost 50% increase in a population segment that currently exhibits half of all cancer incidences.

```{r, echo=FALSE, fig.width=6, fig.height=3}
p_borders_demo
```

# Key Points
- continuing steady rise in cancer incidence and cancer incidence rate 
- three skin cancer sites show notable increases, although these increases are also evident nationwide
- population demographic changes will likely continue to exacerbate the overall increase in cancer incidence


# Sources

Cancer Incidence Data: [https://www.opendata.nhs.scot/dataset/annual-cancer-incidence](https://www.opendata.nhs.scot/dataset/annual-cancer-incidence)

NRS Population Projections: 
[https://www.nrscotland.gov.uk/statistics-and-data/statistics/statistics-by-theme/population/](https://www.nrscotland.gov.uk/statistics-and-data/statistics/statistics-by-theme/population/population-projections/sub-national-population-projections/2016-based/detailed-tables)

Analysis: [https://github.com/jakeybob/cancer_incidence](https://github.com/jakeybob/cancer_incidence)