---
title: "**NHS Borders** -- Cancer Incidence"
author: 
- "John Robert Taylor"
- "[bob.taylor@nhs.net](mailto:bob.taylor@nhs.net)"
- "candidate ID: &nbsp; **20214**"
output:
  pdf_document:
    latex_engine: xelatex
    template: NULL
    keep_tex: false
mainfont: Source Sans Pro
documentclass: proc
urlcolor: blue
geometry:
- layout=a4paper
- top=-10mm
- left=20mm
- right=20mm
- bottom=30mm
subparagraph: yes
header-includes:
- \usepackage{titlesec}
- \titlespacing{\section}{0pt}{2\parskip}{\parskip}
- \titlespacing{\subsection}{0pt}{2\parskip}{\parskip}
- \pagenumbering{gobble}
- \usepackage{titling}
- \setlength{\droptitle}{1.5cm}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Current Incidence Levels
NHS Borders has seen a steady increase in the total cancer incidence (for all cancer types, note some skin cancers are excluded). In 2017, there were 894 incidences of cancer in NHS Border, compared to 702 ten years previously.

```{r, echo=FALSE}
plot(pressure)
```

This trend is mirrored in the rate of total cancer incidence, so is not explicable merely by changing total population numbers.


# Trends and Comparisons

# Future Implications

Future treatment services will increasingly need to be geared towards the need of an ageing population

# Sources

Cancer Incidence Data: [https://www.opendata.nhs.scot/dataset/annual-cancer-incidence](https://www.opendata.nhs.scot/dataset/annual-cancer-incidence)

NRS Population Projections: [https://www.nrscotland.gov.uk/statistics-and-data/statistics/statistics-by-theme/population/](https://www.nrscotland.gov.uk/statistics-and-data/statistics/statistics-by-theme/population/population-projections/sub-national-population-projections/2016-based/detailed-tables)

Analysis: [https://github.com/jakeybob/cancer_incidence](https://github.com/jakeybob/cancer_incidence)